% -*- compile-command: "rubber --unsafe -d LaTeX-checklist.tex" -*-
\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[most]{tcolorbox}
\usepackage{hyperref}
\usepackage{minted}

\title{Hendrix \protect\LaTeX\ checklist for students}

\newtcblisting{good}{colframe=green!50!black}
\newtcblisting{bad}{colframe=red!50!black}

\begin{document}

\maketitle
\def\arraystretch{1.5}

\LaTeX\ is the \emph{de facto} standard system for creating and
distributing high-quality documents in mathematics and computer
science, and you may be required to use it in several classes
throughout the Mathematics and Computer Science majors.

For creating \LaTeX\ documents, we recommend using
\href{http://overleaf.com}{Overleaf}, a free online editor that even
allows multiple people to easily collaborate on the same document.
For learning about the basics of \LaTeX, we recommend working through
some of the exercises in the
\href{https://www.bates.edu/mathematics/resources/latex-manual/}{Bates
  LaTeX Manual}, or the \href{https://overleaf.com/learn/latex}{Overleaf
  tutorials}.

This document is a simple checklist you can use to ensure you are
following some basic best practices for creating professional-looking
\LaTeX\ documents.  It can be found at \url{https://github.com/Hendrix-CS/XXX}.

Throughout this document, examples of \textbf{bad} \LaTeX\ style are typeset in
a red box, like the one below.  You can see the \LaTeX\ code in the top half of
the box, and the resulting typeset document on the bottom.
\begin{bad}
$2^120 is a big number$.  Don't copy the red boxes!
\end{bad}

Examples of good style, on the other hand, are typeset in a green box,
like this:
\begin{good}
$2^{120}$ is a big number.  Pay attention to the green boxes!
\end{good}

\section*{Formatting}
\label{sec:formatting}

\begin{itemize}
\item Use \verb|``| (two back ticks) and \verb|''| (two single quotes)
  to create quotation marks:
\begin{good}
``like this!''
\end{good}
If you use normal double quote charaters, the result will be typeset
using two ``close quote'' characters.
\begin{bad}
"These quote marks look bad."
\end{bad}

\item Use \verb|\emph{...}| to italicize things like Latin abbreviations
  (such as \emph{e.g.} and \emph{i.e.}) or words you are going to
  define.
\begin{good}
A \emph{snoozle} is a special kind of matrix which has all positive entries, \emph{i.e.} all the entries are greater than zero.
\end{good}
\item There are four kinds of horizontal lines:
  \begin{itemize}
  \item A minus sign is created by using a hyphen inside math mode.
    \begin{good}$(-2) - 3$\end{good}
    \begin{bad}$x$ = -4  % A minus sign outside of math mode\end{bad}
  \item A single hyphen is used for hyphenated words.
    \begin{good}helter-skelter\end{good}
  \item A double hyphen creates an ``en-dash'', which is used for
    ranges.
    \begin{good}Pages 23--69 discuss the period 1952--1987.\end{good}
  \item A triple hyphen creates an ``em-dash'', which is used as
    punctuation to signal a break in thought.  The usual typesetting
    style (in the US, at least) is to leave no spaces on either side
    of the em-dash.
    \begin{good}If this were true---which it isn't---then\end{good}
  \end{itemize}
\item Use \verb|itemize| or \verb|enumerate| for lists.
\begin{good}
\begin{itemize}
  \item One thing
  \item Another thing
  \item Last thing
\end{itemize}
\end{good}
  Using \verb|enumerate| instead of \verb|itemize| causes the items to
  be sequentially numbered.

\item If you use a command in the middle of some text, by default it
  will eat any following whitespace:
\begin{bad}
\LaTeX is great
\end{bad}
  Follow the command with a backslash before the space to force an
  explicit space to appear, like this:
\begin{good}
\LaTeX\ is great
\end{good}
\item \LaTeX\ automatically puts extra space at the end of every
  sentence, but it thinks any punctuation is the end of a sentence, so
  abbreviations can have too much space after them.  Use a backslash
  before the space to force a normal space.  (But this is kind of nitpicky
  since it can be hard to even tell the difference!)
\begin{bad}
Dr. Yorgey likes \LaTeX
\end{bad}
\begin{good}
Dr.\ Yorgey likes \LaTeX
\end{good}

\end{itemize}

\section*{Math}
\label{sec:math}

\begin{itemize}
\item Use the \texttt{amsmath} and \texttt{amssymb} packages by
  putting \verb|\usepackage{amsmath}| and \verb|\usepackage{amssymb}|
  in the \emph{preamble} of your document (that is, the part before
  \verb|\begin{document}|).
\item Use math mode appropriately:
  \begin{itemize}
  \item Use dollar signs for inline mathematics (including variable
    names, numbers, equations, formulas, \dots)
    \begin{good}
In this case, $x$ cannot possibly be $-3$, because we assumed that $x > 10$.
    \end{good}
  \item Use \verb|\[ ... \]| for formulas or equations that should be
    centered on their own line.
    \begin{good}
Solving for $x$, we find that \[ x = 2 + y^2 - \pi, \] which is what we wanted to show.
    \end{good}
 \item Use \verb|$ ... $| or \verb|\[ ... \]| for an entire
   equation/formula, instead of building it out of lots of little
   pieces. Use \verb|\text{...}| if you need some normal text in the
     middle of your formula.
   \begin{bad}
$x$ = $\pm 2$ or $(\sqrt{7} - \pi)$
   \end{bad}
   \begin{good}
$x = \pm 2 \text{ or } (\sqrt{7} - \pi)$
   \end{good}
   \begin{good}
\[ f(x) = \begin{cases} x^2 & \text{if } x > 0 \\ x & \text{otherwise} \end{cases} \]
   \end{good}
  \end{itemize}
\item Use \verb|\mathit{...}| for longer variable names in math mode.
  Without this command, \LaTeX\ thinks all the letters represent
  individual variables being multiplied together, and will insert
  extra space between them.  For example, notice how much space is
  between the letters $F$ and $u$ below:
  \begin{bad}
$myFunction(x) \times 2^n$
  \end{bad}
  With the \verb|\mathit{...}| command, \LaTeX\ knows that the entire
  thing is supposed to be one word:
  \begin{good}
$\mathit{myFunction}(x) \times 2^n$
  \end{good}
\item Use commands like \verb|\log|, \verb|\sin|, \verb|\cos|,
  \verb|\lim| appropriately.  Typically such ``operator words'' are
  typeset with upright letters, unlike variables which
  are typeset in italics.  Using the appropriate commands ensures
  these operators are formatted correctly, with appropriate space
  around them.
  \begin{bad}
$n log(n) + lim_{x \to \infty} \frac{cos \theta}{x^2}$
  \end{bad}
  \begin{good}
$n \log(n) + \lim_{x \to \infty} \frac{\cos \theta}{x^2}$
  \end{good}
\item Use \verb|\dots| for ellipses and \verb|\cdots| for centered
  ellipses. Using actual periods looks too squashed:
  \begin{bad}
$\{1, 2, ..., 10\}$
  \end{bad}
  \begin{bad}
$1 + 2 + ... + 10$
  \end{bad}
  The \verb|\dots| and \verb|\cdots| (centered dots) commands produce
  more evenly spaced dots:
  \begin{good}
$\{1, 2, \dots, 10\}$
  \end{good}
  \begin{good}
$1 + 2 + \cdots + 10$
  \end{good}
\item Use \verb|\begin{align*} ... \end{align*}| for multi-step
  derivations.
  \begin{good}
\begin{align*}
f(n) &= x + x^2 + \cdots + x^n \\
     &= x(1 + x + \cdots + x^{n-1}) \\
     &= x \frac{x^n - 1}{x - 1}
\end{align*}
  \end{good}
\item Use \verb|\mathbb{...}| for appropriate symbols.
  \begin{good}
$\mathbb{R} - \mathbb{N}$
  \end{good}
\item Use \verb|{...}| around multi-character exponents, subscripts,
  \emph{etc.}
  \begin{bad}
$2^15$
  \end{bad}
  \begin{good}
$2^{15}$
  \end{good}
\item Use \verb|\left| and \verb|\right| to make sure delimiters are
  big enough.  In this first example, notice how the parentheses are
  not tall enough to enclose everything inside them:
  \begin{bad}
$(\frac{\pi}{\sum_{k=1}^{\infty} \frac{1}{k^2}})$
  \end{bad}
  Putting the \verb|\left| and \verb|\right| commands before the
  parentheses ensures they are big enough:
  \begin{good}
$\left( \frac{\pi}{\sum_{k=1}^{\infty} \frac{1}{k^2}} \right)$
  \end{good}
\end{itemize}

\section*{Code}

\begin{itemize}
\item Use the \texttt{minted}
  package\footnote{\url{https://www.overleaf.com/learn/latex/Code_Highlighting_with_minted}}
  to typeset nicely syntax-highlighted code.  A \verb|verbatim|
  environment is not very flexible and can only produce black and
  white text:
  \begin{bad}
\begin{verbatim}
def facRec(n: int) -> int:
    if n <= 1:
        return n
    else:
        return n * facRec(n-1)
\end{verbatim}
  \end{bad}
The \texttt{minted} package can produce code with syntax highlighting,
and has many other options as well, such as line numbering, including
code from an external file, \emph{etc.}
  \begin{good}
\begin{minted}{python}
def facRec(n: int) -> int:
    if n <= 1:
        return n
    else:
        return n * facRec(n-1)
\end{minted}
  \end{good}
\item Use a contrasting style like a typewriter font for code, names
  of functions, \emph{etc.}
  \begin{bad}When we call facRec(10),\end{bad}
  \begin{good}When we call \texttt{facRec(10)},\end{good}
  Even better, use the \verb|mintinline| command from the
  \texttt{minted} package:
  \begin{good}When we call \mintinline{python}{facRec(10)},\end{good}
\end{itemize}

\section*{Organization}

\begin{itemize}
\item Use \verb|\chapter|, \verb|\section|, \verb|\subsection|,
  \emph{etc.} appropriately to create sections within your document.
\item Use \verb|\label| and \verb|\ref| appropriately to refer to
  numbered sections, theorems, figures, \emph{etc.}
\end{itemize}


\end{document}
